---
title: "Survival analysis in the HN cohort"
author: "Maria Secrier"
date: "09/10/2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "plots/")
```

## Libraries
```{r libs}
library("survminer")
library("survival")
library("knitr")
library("dplyr")
library("forestmodel")
library ("plyr")
```

## Reading data

```{r readdata}
load("dat.hn.plusgroups.RData")
dat.hn <- dat.hn.plusgroups
dat.hn$Age <- sapply(dat.hn$Age..At.Surgery.,
                     function(x) ifelse(x<=40,"25-40",
                                        ifelse(x<=60,"40-60",ifelse(x<=80,"60-80",">80"))))
dat.hn$Age <- factor(dat.hn$Age,
                     levels=c("25-40","40-60","60-80",">80"))
dat.hn[which(dat.hn$Recurrence.YesNo == "NDF"),]$Recurrence.YesNo <- NA
```

## Survival by exhaustion group

```{r survByImmunityGroup, echo=FALSE, fig.width=10, fig.height=8, dev='pdf'}
fit <- survfit(Surv(OS.days, AliveDead) ~ ImmunityGroup, data=dat.hn)
ggsurvplot(fit, linetype="strata",
           #palette = c("#336699", "#9492D0",
            #          "#F98948","#993955"),
           #conf.int = TRUE, 
           pval=TRUE,#legend.labs=c("Hi", "Sex=2"),
           ggtheme = theme_minimal(), risk.table=TRUE, risk.table.y.text.col = TRUE,
          risk.table.height = 0.25, # the height of the risk table
          risk.table.y.text = FALSE,# show bars instead of names in text annotations
                                    # in legend of risk table.
          ncensor.plot = TRUE,      # plot the number of censored subjects at time t
          ncensor.plot.height = 0.25,
          #conf.int.style = "step",  # customize style of confidence intervals
          surv.median.line = "hv")  # add the median survival pointer.
           #legend.labs = 
           # c("Absent", "Weak", "Exhausted", "PC high"))
```


To double check the labelling is right, plotting the survival values by group and also the exhaustion by group:

```{r survBoxplotByExhGroup, echo=FALSE}
ggplot(dat.hn, aes(ImmunityGroup,OS.days))+
  geom_boxplot()

```

## Multivariate Cox PH analysis

```{r multicox, echo=FALSE}
dat.hn.keep <- dat.hn[which(!is.na(dat.hn$ImmunityGroup)),]
#dat.hn.keep$ImmunityGroup <- factor(dat.hn.keep$ImmunityGroup)
covariates <- c(#"productive_clonality","observed_richness",
                "Age","Sex",
                "Stage..Clinical.Simplified",
                "Tstage","Nstage",
                #"Grade",
                "HPV.Status","Smoker",
                "Basic.Site",
                "Recurrence.YesNo",
                "ImmunityGroup")
multiv_formulas <- paste(covariates,
                        collapse="+") 
res.cox <- coxph(as.formula(paste0("Surv(OS.days, AliveDead) ~ 0+",
                                   multiv_formulas)), 
                 data =  dat.hn)
summary(res.cox)
res.cox.x <- summary(res.cox)
#print(x)
p.value.model <- signif(res.cox.x$wald["pvalue"])
p.value.individual<-signif(res.cox.x$coef[,"Pr(>|z|)"])#wald["pvalue"])
wald.test<-signif(res.cox.x$wald["test"], digits=2)
beta<-signif(res.cox.x$coef[,"coef"], digits=2);#coeficient beta
HR <-signif(res.cox.x$coef[,"exp(coef)"], digits=2);#exp(beta)
HR.confint.lower <- signif(res.cox.x$conf.int[,"lower .95"], 2)
HR.confint.upper <- signif(res.cox.x$conf.int[,"upper .95"],2)
#HR <- paste0(HR, " (", 
#             HR.confint.lower, "-", HR.confint.upper, ")")
res.cox.multi<-data.frame(cbind(rownames(res.cox.x$coef),
                                p.value.model,wald.test,
                                p.value.individual,
           beta, HR,HR.confint.lower, HR.confint.upper))
colnames(res.cox.multi) <- c("Covariate","P-value (model)",
                   "Wald test (model)",
                   "P-value (category)","Beta-coefficient",
                   "HR","95% CI for HR (lower)",
                   "95% CI for HR (upper)")
write.csv(res.cox.multi[,1:6],"HNcohort.coxph_OS.multivariate.CTME.csv",
          row.names=FALSE, quote=FALSE)
save(res.cox.multi,file="res.cox.multi.OS.CTME.RData")
```

Multivariate model:

# ```{r multivariateCox.HRs.extended, echo=FALSE, fig.width=10,fig.height=8}
# ggforest(res.cox,data =  dat.hn, main="",fontsize=0.6)
# ```

With added details:

# ```{r multivariateCox.HRs.detailed.extended,echo=FALSE, fig.width=10,fig.height=8}
# forest_model(res.cox)
# ```


## Multivariate Cox PH analysis including most data and no TCR info

```{r multicox2, echo=FALSE}

dat.hn.totest <- dat.hn[which(dat.hn$Recurrence.YesNo %in% c("Yes","No",NA)),]
dat.hn.totest$Nstage.simp <- sapply(dat.hn.totest$Nstage,
                                    function(x) ifelse(x==0,0,1))
dat.hn.totest$Stage <- sapply(dat.hn.totest$Stage..Clinical.Simplified,
                              function(x) ifelse(x %in% c("I","II"),"Early","Late"))
covariatesOrig <- c("Age..At.Surgery.",
  "Sex",
                "Stage","Tstage",
                "Nstage",
                "HPV.Status","Smoker",
                "Basic.Site",
                #"Grade",
                "Recurrence.YesNo",
  "ImmunityGroup")

dat.hn.totest$Basic.Site <- factor(dat.hn.totest$Basic.Site, levels=c("Oral cavity","Larynx","Pharynx","Other"))
#dat.hn.totest$ImmunityGroup <- factor(dat.hn.totest$ImmunityGroup, levels=c("Absent","Weak","Exhausted","PlasmaCellHigh"))
dat.hn.totest$Tstage <- factor(dat.hn.totest$Tstage)
dat.hn.totest$Nstage <- factor(dat.hn.totest$Nstage)

covariates <- c("Age","Sex",
                "ClinicalStage","Tstage",
                "Nstage",
                "HPVstatus","Smoker",
                "TumourLocation",
                "Recurrence",
  "ImmunityGroup")

colnames(dat.hn.totest)[match(covariatesOrig,colnames(dat.hn.totest))] <- 
  covariates


multiv_formulas <- paste(covariates,
                        collapse="+") 
res.cox <- coxph(as.formula(paste0("Surv(OS.days, AliveDead) ~ 0+",
                                   multiv_formulas)), 
                 data =  dat.hn.totest)
summary(res.cox)
res.cox.x <- summary(res.cox)
#print(x)
p.value.model <- signif(res.cox.x$wald["pvalue"])
p.value.individual<-signif(res.cox.x$coef[,"Pr(>|z|)"])#wald["pvalue"])
wald.test<-signif(res.cox.x$wald["test"], digits=2)
beta<-signif(res.cox.x$coef[,"coef"], digits=2);#coeficient beta
HR <-signif(res.cox.x$coef[,"exp(coef)"], digits=2);#exp(beta)
HR.confint.lower <- signif(res.cox.x$conf.int[,"lower .95"], 2)
HR.confint.upper <- signif(res.cox.x$conf.int[,"upper .95"],2)
#HR <- paste0(HR, " (", 
#             HR.confint.lower, "-", HR.confint.upper, ")")
res.cox.multi<-data.frame(cbind(rownames(res.cox.x$coef),
                                p.value.model,wald.test,
                                p.value.individual,
           beta, HR,HR.confint.lower, HR.confint.upper))
colnames(res.cox.multi) <- c("Covariate","P-value (model)",
                   "Wald test (model)",
                   "P-value (category)","Beta-coefficient",
                   "HR","95% CI for HR (lower)",
                   "95% CI for HR (upper)")
write.csv(res.cox.multi[,1:6],"HNcohort.coxph_OS.multivariate.noTCR.CTME.csv",
          row.names=FALSE, quote=FALSE)
save(res.cox.multi,file="res.cox.multi.OS.noTCR.CTME.RData")
```

Multivariate model, no TCR:

# ```{r multivariateCox.HRs.extended.noTCR, echo=FALSE, fig.width=10,fig.height=8, dev='pdf'}
# ggforest(res.cox,data =  dat.hn.totest, main="",fontsize=0.6)
# ```

With added details, no TCR:

# ```{r multivariateCox.HRs.detailed.extended.noTCR,echo=FALSE, fig.width=10,fig.height=8,dev='pdf'}
# forest_model(res.cox)
# ```

Also summarise the full clinical data:

```{r summaryTable, echo=FALSE, dev='pdf'}
library(gtsummary)
load("dat.annot.plusimm.2021.RData")
dat.hn <- dat.annot.plusimm

# make dataset with a few variables to summarize
dat.sum <- dat.hn %>% select(Race, Sex, Stage,HPV.category,Tumor.Site,
                             Smoking.Category)

# summarize the data with our package
table1 <- tbl_summary(dat.sum)
table1
table1 %>%
  as_gt() %>%
  gt::gtsave(filename = "summaryTable.png") 

```